Semantics for the $ \lambda $-calculus

### 2024-02-14 report/chapters/algebraic-structures.tex
Update the definition of fibration, write about the algebra and presheaf fibrations

\begin{definition}[Pullback of algebras]
  If we have a morphism of algebraic theories $ f: T^\prime \to T $, we have a functor $ \Alg T \to \Alg T^\prime $. It endows $ T^\prime $-algebras with an action from $ T $ given by $ g \bullet_{T^\prime} a = f(g) \bullet_T a $. Then $ T^\prime $-algebra morphisms commute with this $ T $-action, so we indeed have a functor.
\end{definition}

\begin{lemma}
  \TODO Fibration
\end{lemma}

@@

\begin{lemma}
  \TODO Fibration
\end{lemma}

@@

\begin{lemma}
  There is an adjoint equivalence between the category $ \Alg{F(S)} $ and the coslice category (\TODO) $ S \downarrow \SET $.
\end{lemma}
\begin{proof}
  For the equivalence, we send a $ F(S) $-algebra $ A $ to the set $ A $ with morphism $ s \mapsto s \bullet () $. An algebra morphism $ f: A \to B $ is sent to the coslice morphism $ f: (S \to A) \to (S \to B) $. This constitutes a functor.

>>>>

\begin{definition}[Pullback of algebras]\label{def:algebra-pullback}
  If we have a morphism of algebraic theories $ f: T^\prime \to T $, we have a functor $ \Alg T \to \Alg T^\prime $. It endows $ T^\prime $-algebras with an action from $ T $ given by $ g \bullet_{T^\prime} a = f(g) \bullet_T a $. Then $ T^\prime $-algebra morphisms commute with this $ T $-action, so we indeed have a functor.
\end{definition}

We can also consider the category of `all' algebraic theory algebras together. That is, the category $ C $ with $ C_0 = \sum_{T : \AlgTh}, \Alg T $ and $ C((T, A), (T^\prime, A^\prime)) $ consisting of pairs $ (f, f^\prime): \AlgTh(T, T^\prime) \times \SET(A, A^\prime) $ such that for all $ t: T_n $ and $ a: A^n $,
\[ f^\prime(t \bullet a) = f(t) \bullet (f^\prime(a_i))_i. \]
We then have a functor $ P: C \to \AlgTh $, projecting onto the first coordinate.

\begin{lemma}\label{lem:algebra-fibration}
  $ P $ is a fibration.
\end{lemma}
\begin{proof}
  Given an algebraic theory morphism $ f: \AlgTh(S, T) $ and a $ T $-algebra $ A_T $, Definition \ref{def:algebra-pullback} gives an $ S $-algebra $ A_S $ with underlying set $ A_T $. The cartesian morphism is $ (f, \id{A_T}): C((S, A_S), (T, A_T)) $.

  It is cartesian because for $ (R, A_R): C $ and $ (g, g^\prime): C((R, A_R), (T, A_T)) $ and $ h: \AlgTh(R, S) $ with $ h \cdot f = g $, the required morphism $ C((R, A_R), (S, A_S)) $ is given by $ g^\prime: \SET(A_R, A_S) $.
\end{proof}

@@

An analogue to Lemma \ref{lem:algebra-fibration} shows that, like with algebras, the total category of presheaves is fibered over the category of algebraic theories.

@@

\begin{lemma}
  There is an adjoint equivalence between the category $ \Alg{F(S)} $ and the coslice category $ S \downarrow \SET $.
\end{lemma}
\begin{proof}
  For the equivalence, we send a $ F(S) $-algebra $ A $ to the set $ A $ with morphism $ s \mapsto s \bullet () $. An algebra morphism $ f: A \to B $ is sent to the coslice morphism $ f: (S \to A) \to (S \to B) $. This constitutes a functor.

### 2024-02-14 report/chapters/category-theory.tex
Update the definition of fibration, write about the algebra and presheaf fibrations

One concept in category theory that can be used to describe a lot of limits and adjunctions is that of a universal arrow (see for example \cite{MacLane}, Part III)
\begin{definition}
  A \iindex{universal arrow} from an object $ c: C $ to a functor $ F: D \to C $ consists of an object $ d: D $ and a morphism $ f: D(c, F(d)) $ such that for every similar pair $ (d^\prime, f^\prime) $, $ f^\prime $ factors uniquely as $ f \cdot F(g) $ for some $ g: C(d, d^\prime) $:
  \begin{center}
    \begin{tikzcd}
      c \arrow[d, "f"'] \arrow[rd, "f^\prime"] &\\
      F(d) \arrow[r, "F(g)"'] & F(d^\prime)
    \end{tikzcd}
  \end{center}
\end{definition}

There is also the dual concept of a universal arrow $ (d, f) $ from a functor $ F $ to an object $ c: C $. Its universal property yields the following diagram:
\begin{center}
  \begin{tikzcd}
    F(d^\prime)\arrow[rd, "f^\prime"']  \arrow[r, "F(g)"] & F(d) \arrow[d, "f"]\\
    & c
  \end{tikzcd}
\end{center}

@@

\begin{definition}
  A morphism $ f: E(y, z) $ is called \iindex{cartesian} if for all $ g: E(x, z) $ and $ w: B(P(x), P(y)) $ with $ w \cdot P(f) = P(g) $, there exists $ \bar w: E(x, y) $ such that $ P(\bar w) = w $ and $ \bar w \cdot f = g $, like illustrated in the following diagram from \cite{nlab:grothendieck_fibration}
  \begin{center}
    \begin{tikzcd}[row sep = .5in]
      E \arrow[d, "P"] &x \arrow[rr, "\forall g", bend left] \arrow[r, "\exists! \bar w"'] & y \arrow[r, "f"'] & z\\
      B & P(x) \arrow[r, "\forall w"'] \arrow[rr, "P(g)", bend left] & P(y) \arrow[r, "P(f)"'] & P(z)
    \end{tikzcd}
  \end{center}
\end{definition}

\begin{definition}
  $ P $ is a \iindex{fibration} if for all $ y: E $ and morphisms $ f: B(x, P(y)) $, there exists a cartesian morphism $ \bar f: E(\bar x, y) $ such that $ P(\bar f) = f $.
\end{definition}

\section{(Co)slice categories}
Given an object in a category $ c: C $, the morphisms to and from $ c $ constitute the slice and coslice categories
\begin{definition}
  The \iindex{slice category} $ C \downarrow c $ is the category with as objects the morphisms to $ c $:
  \[ (C \downarrow c)_0 = \sum_{c^\prime: C} C(c^\prime, c). \]
  The morphisms from $ (c^\prime, f) $ to $ (c^{\prime\prime}, f^\prime) $ are the morphisms $ g: c^\prime \to c^{\prime\prime} $ making the following diagram commute.
  \begin{center}
    \begin{tikzcd}[column sep=.5cm]
      c^\prime \arrow[rr, "g"] \arrow[rd, "f"'] & & c^{\prime\prime} \arrow[ld, "f^\prime"]\\
      & c
    \end{tikzcd}
  \end{center}
\end{definition}
The \textit{coslice category}\index{slice category!co-} $ c \downarrow C $ is similar, but with the morphisms from $ c $:
\[ (c \downarrow C)_0 = \sum_{c^\prime: C} C(c, c^\prime). \]

@@

\begin{definition}
  Given an object $ c $ in a category $ C $ with terminal object $ t $. A \iindex{global element} of $ c $ is a morphism $ C(t, c) $.
\end{definition}

>>>>

One concept in category theory that can be used to describe a lot of limits and adjunctions is that of a universal arrow (see for example \cite{MacLane}, Part III)
\begin{definition}
  A \iindex{universal arrow} from an object $ c: C $ to a functor $ F: D \to C $ consists of an object $ d: D $ and a morphism $ f: D(c, F(d)) $ such that for every similar pair $ (d^\prime, f^\prime) $, $ f^\prime $ factors uniquely as $ f \cdot F(g) $ for some $ g: C(d, d^\prime) $:
  \begin{center}
    \begin{tikzcd}
      c \arrow[d, "f"'] \arrow[rd, "f^\prime"] &\\
      F(d) \arrow[r, "F(g)"', dashed] & F(d^\prime)
    \end{tikzcd}
  \end{center}
\end{definition}

There is also the dual concept of a universal arrow $ (d, f) $ from a functor $ F $ to an object $ c: C $. Its universal property yields the following diagram:
\begin{center}
  \begin{tikzcd}
    F(d^\prime)\arrow[rd, "f^\prime"']  \arrow[r, "F(g)", dashed] & F(d) \arrow[d, "f"]\\
    & c
  \end{tikzcd}
\end{center}

@@

\begin{definition}
  A morphism $ f: E(y, z) $ is called \iindex{cartesian} if for all $ g: E(x, z) $ and $ h: B(P(x), P(y)) $ with $ h \cdot P(f) = P(g) $, there exists $ \bar h: E(x, y) $ such that $ P(\bar h) = h $ and $ \bar h \cdot f = g $, like illustrated in the following diagram from \cite{nlab:grothendieck_fibration}
  \begin{center}
    \begin{tikzcd}[row sep = .5in]
      E \arrow[d, "P"] &x \arrow[rr, "\forall g", bend left] \arrow[r, "\exists! \bar h"', dashed] & y \arrow[r, "f"'] & z\\
      B & P(x) \arrow[r, "\forall h"'] \arrow[rr, "P(g)", bend left] & P(y) \arrow[r, "P(f)"'] & P(z)
    \end{tikzcd}
  \end{center}
\end{definition}

\begin{definition}
  $ P $ is a \iindex{fibration} if for all $ y: E $ and morphisms $ f: B(x, P(y)) $, there exist an object $ \bar x: E $ and a cartesian morphism $ \bar f: E(\bar x, y) $ such that $ P(\bar x) = x $ and $ P(\bar f) = f $:
  \begin{center}
    \begin{tikzcd}
      E \arrow[d, "P"] & \bar x \arrow[r, "\exists \bar f", dashed] & y\\
      B & x \arrow[r, "\forall f"] & P(y)
    \end{tikzcd}
  \end{center}
\end{definition}

\section{(Co)slice categories}
Given an object in a category $ c: C $, the morphisms to and from $ c $ constitute the slice and coslice categories
\begin{definition}
  The \iindex{slice category} $ C \downarrow c $ is the category with as objects the morphisms to $ c $:
  \[ (C \downarrow c)_0 = \sum_{c^\prime: C} C(c^\prime, c). \]
  The morphisms from $ (c^\prime, f) $ to $ (c^{\prime\prime}, f^\prime) $ are the morphisms $ g: c^\prime \to c^{\prime\prime} $ making the following diagram commute.
  \begin{center}
    \begin{tikzcd}[column sep=.5cm]
      c^\prime \arrow[rr, "g"] \arrow[rd, "f"'] & & c^{\prime\prime} \arrow[ld, "f^\prime"]\\
      & c
    \end{tikzcd}
  \end{center}
\end{definition}
The \textit{coslice category}\index{slice category!co-} $ c \downarrow C $ is similar, but with the morphisms \textit{from} $ c $ instead of \textit{to} $ c $:
\[ (c \downarrow C)_0 = \sum_{c^\prime: C} C(c, c^\prime). \]

@@

\begin{definition}
  Let $ C $ be a category with terminal object $ t $. For an object $ c: C $, a \iindex{global element} of $ c $ is a morphism $ f: C(t, c) $.
\end{definition}
